---
title: "Forecasting Competition"
author: "Saemin Kim and Matthew Fischer"
output: pdf_document
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

# Data Wrangling and Vanila Model

## Setting R code chunk options

First R code chunk is used for setting the options for all R code chunks. The choice echo=TRUE means both code and output will appear on report, include = FALSE neither code nor output is printed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

## Loading packages and initializing

Second R code chunk is for loading packages. By setting message = FALSE, the code will appear but not the output.

```{r package, message=FALSE}
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)

library(kableExtra)
library(readxl)
library(openxlsx)
```

## Importing data

```{r}
data <- read_excel("load.xlsx")

daily_data <- data %>%
  mutate(
    date = as.Date(date),
    Year = year(date),
    Month = month(date),
    Day = day(date)
  ) %>%
  
  rowwise() %>%
  mutate(
    daily_avg = mean(c_across(h1:h24), na.rm = TRUE)
  ) %>%
  ungroup() %>%
  
  select(meter_id, date, Year, Month, Day, daily_avg)

daily_data <- daily_data %>%
  filter(!is.na(daily_avg), !is.nan(daily_avg), is.finite(daily_avg))
```

```{r}
ggplot(daily_data, aes(x = date, y = daily_avg)) +
  geom_line() +
  labs(title = "Daily Average Load", y = "Daily Average", x = "Date")
```

```{r}
data_ts <- msts(daily_data$daily_avg, 
                            seasonal.periods =c(7, 365),
                            start=c(2005,01))

SNAIVE_data <- snaive(data_ts, h=31)
plot(SNAIVE_data)

forecast_values <- as.numeric(SNAIVE_data$mean)
```

```{r}
submission <- read.csv("submission_template.csv")

submission$load <- forecast_values
write.csv(submission, "seasonal_naive_model.csv", row.names = FALSE)
```

# External Variables and Train/Test Split

## Temperature, Humidity Data Wrangling

```{r external_data, message=FALSE, warning=FALSE}

temp <- read_excel("temperature.xlsx")
humid <- read_excel("relative_humidity.xlsx")

# daliy avg of temp
temp_daily <- temp %>%
  mutate(date = as.Date(date)) %>%
  rowwise() %>%
  # hourly avg of all t_ws1 ~ t_ws28
  mutate(hourly_temp_avg = mean(c_across(starts_with("t_ws")), na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(date) %>%
  # daily avg
  summarise(daily_temp = mean(hourly_temp_avg, na.rm = TRUE))

# daliy avg of humid 
humid_daily <- humid %>%
  mutate(date = as.Date(date)) %>%
  rowwise() %>%
  # hourly avg of all rh_ws1 ~ rh_ws28
  mutate(hourly_humid_avg = mean(c_across(starts_with("rh_ws")), na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(date) %>%
  # daily avg
  summarise(daily_humid = mean(hourly_humid_avg, na.rm = TRUE))

# Join the data
daily_data <- daily_data %>%
  left_join(temp_daily, by = "date") %>%
  left_join(humid_daily, by = "date")

head(daily_data)
```

## Train and Validation Split

```{r train_validation_split, message=FALSE, warning=FALSE}
# Training = 2005-01-01 to 2010-06-30 / Validation = 2010-07-01 to 2010-07-31
train_data <- daily_data %>% filter(date >= as.Date("2005-01-01") & date <= as.Date("2010-06-30"))
valid_data <- daily_data %>% filter(date >= as.Date("2010-07-01") & date <= as.Date("2010-07-31"))

# Create msts (Multi-Seasonal Time Series) objects for training and validation
# Apply weekly (7) and yearly (365) seasonal periods
train_ts <- msts(train_data$daily_avg, seasonal.periods = c(7, 365), start = c(2005, 1))
valid_ts <- msts(valid_data$daily_avg, seasonal.periods = c(7, 365), start = c(2010, 182)) 
# 182 represents the approximate day of the year for July 1st
```

# Forecast Models

```{r final_kaggle_submission, message=FALSE, warning=FALSE}
# 1. Create msts object using the FULL dataset (Jan 2005 - Jun 2011)
full_ts <- msts(daily_data$daily_avg, seasonal.periods = c(7, 365), start = c(2005, 1))

# --- Model 1: TBATS Final Forecast ---
print("Retraining TBATS on full data...")
fit_tbats_full <- tbats(full_ts)
forecast_tbats_full <- forecast(fit_tbats_full, h = 31)

# Save TBATS prediction to CSV
submission$load <- as.numeric(forecast_tbats_full$mean)
write.csv(submission, "model 1 submission.csv", row.names = FALSE)
print("Saved model_1_submission.csv")
```

