---
title: "Forecasting Competition"
author: "Saemin Kim and Matthew Fischer"
output: pdf_document
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

# 1. Data Wrangling and Vanila Model

## Setting R code chunk options

First R code chunk is used for setting the options for all R code chunks. The choice echo=TRUE means both code and output will appear on report, include = FALSE neither code nor output is printed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

## Loading packages and initializing

Second R code chunk is for loading packages. By setting message = FALSE, the code will appear but not the output.

```{r package, message=FALSE}
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)

library(kableExtra)
library(readxl)
library(openxlsx)
```

## Importing data

```{r}
data <- read_excel("load.xlsx")

daily_data <- data %>%
  mutate(
    date = as.Date(date),
    Year = year(date),
    Month = month(date),
    Day = day(date)
  ) %>%
  
  rowwise() %>%
  mutate(
    daily_avg = mean(c_across(h1:h24), na.rm = TRUE)
  ) %>%
  ungroup() %>%
  
  select(meter_id, date, Year, Month, Day, daily_avg)

daily_data <- daily_data %>%
  filter(!is.na(daily_avg), !is.nan(daily_avg), is.finite(daily_avg))
```

```{r}
ggplot(daily_data, aes(x = date, y = daily_avg)) +
  geom_line() +
  labs(title = "Daily Average Load", y = "Daily Average", x = "Date")
```

```{r}
data_ts <- msts(daily_data$daily_avg, 
                            seasonal.periods =c(7, 365),
                            start=c(2005,01))

SNAIVE_data <- snaive(data_ts, h=31)
plot(SNAIVE_data)

forecast_values <- as.numeric(SNAIVE_data$mean)
```

```{r}
submission <- read.csv("submission_template.csv")

submission$load <- forecast_values
write.csv(submission, "seasonal_naive_model.csv", row.names = FALSE)
```

# 2. External Variables and Train/Test Split

## Temperature, Humidity Data Wrangling

```{r external_data, message=FALSE, warning=FALSE}

temp <- read_excel("temperature.xlsx")
humid <- read_excel("relative_humidity.xlsx")

# daliy avg of temp
temp_daily <- temp %>%
  mutate(date = as.Date(date)) %>%
  rowwise() %>%
  # hourly avg of all t_ws1 ~ t_ws28
  mutate(hourly_temp_avg = mean(c_across(starts_with("t_ws")), na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(date) %>%
  # daily avg
  summarise(daily_temp = mean(hourly_temp_avg, na.rm = TRUE))

# daliy avg of humid 
humid_daily <- humid %>%
  mutate(date = as.Date(date)) %>%
  rowwise() %>%
  # hourly avg of all rh_ws1 ~ rh_ws28
  mutate(hourly_humid_avg = mean(c_across(starts_with("rh_ws")), na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(date) %>%
  # daily avg
  summarise(daily_humid = mean(hourly_humid_avg, na.rm = TRUE))

# Join the data
daily_data <- daily_data %>%
  left_join(temp_daily, by = "date") %>%
  left_join(humid_daily, by = "date")

head(daily_data)
```

## Train and Validation Split

```{r train_validation_split, message=FALSE, warning=FALSE}
# Training = 2005-01-01 to 2010-06-30 / Validation = 2010-07-01 to 2010-07-31
train_data <- daily_data %>% filter(date >= as.Date("2005-01-01") & date <= as.Date("2010-06-30"))
valid_data <- daily_data %>% filter(date >= as.Date("2010-07-01") & date <= as.Date("2010-07-31"))

# Create msts (Multi-Seasonal Time Series) objects for training and validation
# Apply weekly (7) and yearly (365) seasonal periods
train_ts <- msts(train_data$daily_avg, seasonal.periods = c(7, 365), start = c(2005, 1))
valid_ts <- msts(valid_data$daily_avg, seasonal.periods = c(7, 365), start = c(2010, 182)) 
# 182 represents the approximate day of the year for July 1st
```

# 3. Top 5 Models: Analysis and Model Selection

In the comprehensive modeling phase of this project, we evaluated 19 distinct time series forecasting approaches, ranging from traditional statistical methods (TBATS, STLF-ARIMA) to machine learning techniques (NNAR, Prophet). Across all experiments, the Facebook Prophet framework consistently demonstrated superior predictive accuracy for the daily electricity load, ultimately securing all Top 5 positions (Please refer to the appendix for the score summary board and the complete code and performance evaluation of all models.)

Our baseline Prophet model (Model 8, Rank 5) provided a strong starting point, but we observed immediate improvements by incorporating deterministic calendar events, specifically US Holidays (Model 11, Rank 4). Since industrial and commercial electricity consumption drops significantly during holidays like Independence Day, allowing the model to account for these anomalies inherently reduced forecasting errors. From there, applying targeted hyperparameter tuning further optimized our predictions (Models 12, 13, and 14, tied for Rank 3). Specifically, we increased the flexibility of the yearly seasonality (yearly.seasonality = 15) and adjusted the trend flexibility by setting changepoint.prior.scale = 0.1. This adjustment allowed the model to adapt more rapidly to recent trend shifts in the load data.

The most critical breakthrough in our modeling strategy was the transition from additive to multiplicative seasonality. In electricity demand, the amplitude of seasonal fluctuations (especially during peak summer cooling periods) tends to grow proportionally with the overall trend. By changing the seasonality mode to multiplicative (Model 16, Rank 2) and combining it with conditional weekly seasonality to separate summer and non-summer behavioral patterns, we achieved our best-performing model: Model 17 (Rank 1, Score: 0.0963).

Interestingly, our experiments revealed that incorporating external weather regressors (temperature and humidity forecasts) degraded model performance (e.g., Models 18 and 19). Because the weather variables for July 2011 had to be forecasted themselves, they introduced additional uncertainty and noise into the final load prediction. Consequently, our optimal strategy relied strictly on robust historical load patterns, explicit holiday effects, and dynamic seasonal structures without external weather dependencies.

# APPENDIX. Forecast Models and Scores

## Score Summary

| Model Number | Model Description | Score on Kaggle | Rank |
|:-----------------|:-----------------|:-----------------|:-----------------|
| **Model 1** | TBATS | 0.1021 |  |
| **Model 2** | STLF-ARIMA | 0.1315 |  |
| **Model 3** | STLF + ETS | 0.1330 |  |
| **Model 4-1** | ARIMA + Fourier terms (K = c(3,10)) | 0.1350 |  |
| **Model 4-2** | ARIMA + Fourier terms (K = c(3,15)) | 1.4840 |  |
| **Model 4-3** | ARIMA + Fourier terms (K = c(2,5)) | 0.1469 |  |
| **Model 5** | NNAR | 0.1770 |  |
| **Model 6** | Ensemble Model (Averaging best outputs) | 0.1211 |  |
| **Model 7** | ARIMAX with temperature forecast | 0.2172 |  |
| **Model 8** | Prophet | 0.0973 |  **TOP 5** |
| **Model 9** | Prophet + Temp data | 0.1052 |  |
| **Model 10** | Prophet + ARIMAX | 0.1488 |  |
| **Model 11** | Prophet + US Holiday | 0.0971 |  **TOP 4** |
| **Model 12** | Prophet + US Holiday + Parameter tuning | 0.0966 |  **TOP 3** |
| **Model 13** | Weighted Ensemble (Tuned Prophet + Base Prophet) | 0.0966 |  **TOP 3** |
| **Model 14** | Prophet + US Holiday + Conditional Seasonality | 0.0966 |  **TOP 3** |
| **Model 15** | NNAR + Temperature | 0.1287 |  |
| **Model 16** | Prophet + US Holiday + Multiplicative Seasonality | 0.0964 |  **TOP 2** |
| **Model 17** | Prophet + US Holiday + Multiplicative + Conditional Seasonality | **0.0963** |  **TOP 1** |
| **Model 18** | Prophet + Holiday + Multiplicative + Temp + Humid | 0.1057 |  |
| **Model 19** | Prophet + Holiday + Multi + Cond + Temp + Humid | 0.1063 |  |

## Model 1 : TBATS

Score on Kaggle: 0.1021

```{r model1, message=FALSE, warning=FALSE}
# Create msts object using the FULL dataset (Jan 2005 - Jun 2011)
full_ts <- msts(daily_data$daily_avg, seasonal.periods = c(7, 365), start = c(2005, 1))

# --- Model 1: TBATS Final Forecast ---
print("Retraining TBATS on full data...")
fit_tbats_full <- tbats(full_ts)
forecast_tbats_full <- forecast(fit_tbats_full, h = 31)

# Save TBATS prediction to CSV
submission$load <- as.numeric(forecast_tbats_full$mean)
write.csv(submission, "model 1 submission.csv", row.names = FALSE)
print("Saved model 1 submission.csv")
```

## Model 2: STLF-ARIMA

Score on Kaggle: 0.1315

```{r model2, message=FALSE, warning=FALSE}
# --- Model 2: STLF (with ARIMA) Final Forecast ---
forecast_arima_full <- stlf(full_ts, method = "arima", h = 31)

# Save STLF-ARIMA prediction to CSV
submission$load <- as.numeric(forecast_arima_full$mean)
write.csv(submission, "model 2 submission.csv", row.names = FALSE)
print("Saved model 2 submission.csv")
```

## Model 3: STLF + ETS

Score on Kaggle: 0.1330

```{r model 3, message=FALSE, warning=FALSE}
# --- Model 3: STLF with ETS (Exponential Smoothing) ---
forecast_ets_full <- stlf(full_ts, method = "ets", h = 31)

submission$load <- as.numeric(forecast_ets_full$mean)
write.csv(submission, "model 3 submission.csv", row.names = FALSE)
print("Saved model 3 submission.csv")
```

## Model 4-1: ARIMA + Fourier terms using K = c(3,10)

Score on Kaggle: 0.1350

```{r}
# --- Model 4: Harmonic Regression (ARIMA with Fourier terms) ---
# Using K = c(3, 10)
fit_fourier <- auto.arima(full_ts, seasonal = FALSE, xreg = fourier(full_ts, K = c(3, 10)))
forecast_fourier <- forecast(fit_fourier, xreg = fourier(full_ts, K = c(3, 10), h = 31))

submission$load <- as.numeric(forecast_fourier$mean)
write.csv(submission, "model 4-1 submission.csv", row.names = FALSE)
print("Saved model 4-1 submission.csv")
```

## Model 4-2: ARIMA + Fourier terms using K = c(3,15)

Score on Kaggle: 1.484

```{r}
# --- Model 4: Harmonic Regression (ARIMA with Fourier terms) ---
# Using K = c(3, 15)
fit_fourier <- auto.arima(full_ts, seasonal = FALSE, xreg = fourier(full_ts, K = c(3, 15)))
forecast_fourier <- forecast(fit_fourier, xreg = fourier(full_ts, K = c(3, 15), h = 31))

submission$load <- as.numeric(forecast_fourier$mean)
write.csv(submission, "model 4-2 submission.csv", row.names = FALSE)
print("Saved model 4-2 submission.csv")
```

## Model 4-3: ARIMA + Fourier terms using K = c(2,5)

Score on Kaggle: 0.1469

```{r}
# --- Model 4: Harmonic Regression (ARIMA with Fourier terms) ---
# Using K = c(2, 5)
fit_fourier <- auto.arima(full_ts, seasonal = FALSE, xreg = fourier(full_ts, K = c(2, 5)))
forecast_fourier <- forecast(fit_fourier, xreg = fourier(full_ts, K = c(2, 5), h = 31))

submission$load <- as.numeric(forecast_fourier$mean)
write.csv(submission, "model 4-3 submission.csv", row.names = FALSE)
print("Saved model 4-3 submission.csv")
```

## Model 5: NNAR

Score on Kaggle: 0.1770

```{r}
# --- Model 5: NNAR (Neural Network Time Series Forecast) ---
fit_nnar <- nnetar(full_ts)
forecast_nnar <- forecast(fit_nnar, h = 31)

submission$load <- as.numeric(forecast_nnar$mean)
write.csv(submission, "model 5 submission.csv", row.names = FALSE)
print("Saved model 5 submission.csv")
```

## Model 6: Ensemble Model (Averaging the best outputs)

Score on Kaggle: 0.1211

```{r}
# Averaging the predictions of TBATS, STLF(ARIMA), and STLF(ETS)
ensemble_mean <- (as.numeric(forecast_tbats_full$mean) +
                  as.numeric(forecast_arima_full$mean) + 
                  as.numeric(forecast_ets_full$mean)) / 3

submission$load <- ensemble_mean
write.csv(submission, "model 6 submission.csv", row.names = FALSE)
print("Saved model 6 submission.csv")
```

## Model 7: ARIMAX with temperature forecast

Score on Kaggle: 0.2172

```{r arimax_with_temp_forecast, message=FALSE, warning=FALSE}

# Create a time series object for the historical daily temperature
# Using 365 as the seasonal period since temperature has a strong yearly cycle
temp_ts <- msts(daily_data$daily_temp, seasonal.periods = 365, start = c(2005, 1))

# Fit an ARIMA model to the temperature data
fit_temp <- auto.arima(temp_ts)
# Forecast the temperature for the next 31 days (July 2011)
forecast_temp <- forecast(fit_temp, h = 31)
# Extract the forecasted temperature values as a numeric vector
future_temp <- as.numeric(forecast_temp$mean)


# Forecast the electricity load using the predicted temperatures (ARIMAX)
# Convert full_ts to a simple weekly time series to prevent the auto.arima multi-seasonality error.
# The 'daily_temp' external regressor will naturally capture the yearly (365-day) seasonality.
full_ts_weekly <- ts(daily_data$daily_avg, frequency = 7)

# Fit an ARIMAX model using the weekly ts and historical temperature as an external regressor (xreg)
fit_arimax_full <- auto.arima(full_ts_weekly, xreg = daily_data$daily_temp)

# Forecast the load for July 2011 using the predicted July temperatures as the future xreg
forecast_arimax_final <- forecast(fit_arimax_full, xreg = future_temp, h = 31)

# Save the final predictions to a CSV file for Kaggle submission
submission <- read.csv("submission_template.csv")
submission$load <- as.numeric(forecast_arimax_final$mean)
write.csv(submission, "model 7 submission.csv", row.names = FALSE)
print("Saved model 7 submission.csv")
```

## Model 8: Prophet

Score on Kaggle: 0.0973

```{r prophet_model, message=FALSE, warning=FALSE}
# Load the prophet library
library(prophet)

# 1. Prepare the dataframe in the specific format required by Prophet
df_prophet <- data.frame(ds = daily_data$date, y = daily_data$daily_avg)

# 2. Initialize and fit the Prophet model
# We enable yearly and weekly seasonality. 
# daily.seasonality is set to FALSE because we are working with daily aggregated data.
m <- prophet(yearly.seasonality = TRUE, 
             weekly.seasonality = TRUE, 
             daily.seasonality = FALSE)

m <- fit.prophet(m, df_prophet)

# 3. Create a dataframe for future dates (Next 31 days for July 2011)
future_dates <- make_future_dataframe(m, periods = 31)

# 4. Generate the forecast
forecast_prophet <- predict(m, future_dates)

# Extract only the forecasted values ('yhat') for the last 31 days (July 2011)
prophet_predictions <- tail(forecast_prophet$yhat, 31)

# 5. Save the final predictions
submission <- read.csv("submission_template.csv")
submission$load <- prophet_predictions
write.csv(submission, "model 8 submission.csv", row.names = FALSE)

print("Saved model 8 submission.csv")
```

```{r prophet_visualization, message=FALSE, warning=FALSE}
library(ggplot2)

# 1. Plot the main forecast with historical data and confidence intervals
# Black dots: Actual historical load
# Blue line: Prophet's predicted load
# Light blue shaded area: Uncertainty intervals (Confidence intervals)
print("Plotting the main Prophet forecast...")
plot(m, forecast_prophet) + 
  ggtitle("Prophet Forecast for Daily Load") +
  xlab("Date") + ylab("Daily Average Load")

# 2. Plot the individual components (Trend, Weekly, and Yearly seasonality)
prophet_plot_components(m, forecast_prophet)
```

## Model 9: Prophet + Temp data

Score on Kaggle: 0.1052

```{r prophet_with_regressor, message=FALSE, warning=FALSE}
library(prophet)

print("Preparing data for Prophet with Temperature...")
# 1. Prepare data with both target (y) and the external regressor (temp)
df_prophet_reg <- data.frame(
  ds = daily_data$date, 
  y = daily_data$daily_avg,
  temp = daily_data$daily_temp # Adding historical temperature
)

# 2. Initialize Prophet and ADD the regressor BEFORE fitting
m_reg <- prophet(yearly.seasonality = TRUE, 
                 weekly.seasonality = TRUE, 
                 daily.seasonality = FALSE)
m_reg <- add_regressor(m_reg, 'temp')

print("Fitting Prophet Model with Temperature Regressor...")
m_reg <- fit.prophet(m_reg, df_prophet_reg)

# 3. Create future dataframe
future_dates_reg <- make_future_dataframe(m_reg, periods = 31)
all_temp <- c(daily_data$daily_temp, future_temp)
future_dates_reg$temp <- all_temp

# 5. Generate the final forecast
forecast_prophet_reg <- predict(m_reg, future_dates_reg)

# 6. Extract predictions and save for Kaggle
prophet_reg_predictions <- tail(forecast_prophet_reg$yhat, 31)

submission$load <- prophet_reg_predictions
write.csv(submission, "model 9 submission.csv", row.names = FALSE)
print("Saved model 9 submission.csv")
```

## Model 10: Prophet + ARIMAX

Score on Kaggle: 0.1488

```{r}
ensemble_ultimate <- (prophet_reg_predictions +       
                      as.numeric(forecast_arimax_final$mean)) / 2

submission$load <- ensemble_ultimate
write.csv(submission, "model 10 submission.csv", row.names = FALSE)
print("Saved model 10 submission.csv")
```

## Model 11: Prophet + US Holiday

Score on Kaggle: 0.0971

```{r prophet_with_holidays_only, message=FALSE, warning=FALSE}

# 1. Prepare data
df_prophet_holidays <- data.frame(
  ds = daily_data$date, 
  y = daily_data$daily_avg
)

# 2. Initialize Prophet
m_holidays <- prophet(yearly.seasonality = TRUE, 
                      weekly.seasonality = TRUE, 
                      daily.seasonality = FALSE)

# Add US country holidays (e.g., Independence Day on July 4th)
m_holidays <- add_country_holidays(m_holidays, country_name = 'US')
m_holidays <- fit.prophet(m_holidays, df_prophet_holidays)


# 3. Create future dataframe for July 2011 (31 days)
future_dates_holidays <- make_future_dataframe(m_holidays, periods = 31)

# 4. Generate the final forecast
print("Forecasting with Holidays...")
forecast_prophet_holidays <- predict(m_holidays, future_dates_holidays)

# 5. Extract predictions
prophet_holidays_predictions <- tail(forecast_prophet_holidays$yhat, 31)

submission$load <- prophet_holidays_predictions
write.csv(submission, "model 11 submission.csv", row.names = FALSE)

print("Saved model 11 submission.csv")
```

## Model 12: Prophet + US Holiday + Parameter tuning

Score on Kaggle: 0.0966

```{r prophet_tuned, message=FALSE, warning=FALSE}
# 1. Initialize Prophet with custom hyperparameters
# Increased yearly seasonality flexibility (15 instead of default 10)
# changepoint.prior.scale = 0.1 allows the trend to adapt faster
m_tuned <- prophet(yearly.seasonality = 15, 
                   weekly.seasonality = TRUE, 
                   daily.seasonality = FALSE,
                   changepoint.prior.scale = 0.1)

# Add US holidays
m_tuned <- add_country_holidays(m_tuned, country_name = 'US')

# 2. Fit and Predict
m_tuned <- fit.prophet(m_tuned, df_prophet_holidays)
future_tuned <- make_future_dataframe(m_tuned, periods = 31)
forecast_tuned <- predict(m_tuned, future_tuned)

# 3. Extract and Save
prophet_tuned_preds <- tail(forecast_tuned$yhat, 31)

submission$load <- prophet_tuned_preds
write.csv(submission, "model 12 submission.csv", row.names = FALSE)

print("Saved model 12 submission.csv")
```

## Model 13: Weighted Ensemble (Tuned Prophet + Base Prophet)

Score on Kaggle: 0.0966

```{r final_weighted_ensemble, message=FALSE, warning=FALSE}
# Model A (Rank 1): Tuned Prophet + Holidays (70% weight)
weight_best <- 0.70

# Model B (Rank 2): Base Prophet + Holidays (30% weight)
weight_second <- 0.30

# Calculate the weighted average of the predictions
weighted_ensemble_preds <- (weight_best * prophet_tuned_preds) + 
                           (weight_second * prophet_holidays_predictions)

# Load the submission template and save the final ensemble forecast
submission$load <- weighted_ensemble_preds
write.csv(submission, "model 13 submission.csv", row.names = FALSE)
print("Saved model 13 submission.csv")
```

## Model 14: Prophet + US Holiday + Conditional Seasonality

Score on Kaggle: 0.0966

```{r prophet_conditional, message=FALSE, warning=FALSE}
library(lubridate)

# 1. Prepare data and add conditional columns for Summer vs. Other seasons
df_prophet_ultimate <- data.frame(
  ds = daily_data$date, 
  y = daily_data$daily_avg
)

# Define 'Summer' as June, July, August, and September
df_prophet_ultimate$is_summer <- month(df_prophet_ultimate$ds) %in% c(6, 7, 8, 9)
df_prophet_ultimate$is_other <- !df_prophet_ultimate$is_summer

# 2. Initialize the Prophet model with Tuned Hyperparameters
# Turn OFF default weekly seasonality because we will define custom conditional ones.
m_ultimate <- prophet(yearly.seasonality = 15,          # Tuned parameter
                      weekly.seasonality = FALSE,       # Disabled for custom setup
                      daily.seasonality = FALSE,
                      changepoint.prior.scale = 0.1)    # Tuned parameter

# Add US country holidays
m_ultimate <- add_country_holidays(m_ultimate, country_name = 'US')

# 3. Add Custom Conditional Seasonalities
# This allows the model to learn a different weekly pattern for summer vs other months.
m_ultimate <- add_seasonality(m_ultimate, name = 'summer_weekly', period = 7, 
                              fourier.order = 3, condition.name = 'is_summer')
m_ultimate <- add_seasonality(m_ultimate, name = 'other_weekly', period = 7, 
                              fourier.order = 3, condition.name = 'is_other')
m_ultimate <- fit.prophet(m_ultimate, df_prophet_ultimate)


# 4. Create future dataframe for July 2011 (31 days) and apply conditions
future_ultimate <- make_future_dataframe(m_ultimate, periods = 31)
future_ultimate$is_summer <- month(future_ultimate$ds) %in% c(6, 7, 8, 9)
future_ultimate$is_other <- !future_ultimate$is_summer

# 5. Generate the final forecast
print("Forecasting with Ultimate Prophet...")
forecast_ultimate <- predict(m_ultimate, future_ultimate)

# 6. Extract predictions and save to CSV
prophet_ultimate_preds <- tail(forecast_ultimate$yhat, 31)
submission$load <- prophet_ultimate_preds
write.csv(submission, "model 14 submission.csv", row.names = FALSE)

print("Saved model 14 submission.csv")
```

## Model 15: NNAR + Temperature

Score on Kaggle: 0.1287

```{r nnar_with_temperature, message=FALSE, warning=FALSE}
# Load necessary library
library(forecast)

# 1. Convert historical load to a simple time series object (Weekly frequency)
ts_load <- ts(daily_data$daily_avg, frequency = 7)

# 2. Fit the Neural Network Auto-Regression (NNAR) model
# We use historical daily temperature as an external regressor (xreg)

# The set.seed ensures the neural network initializes the same way every time
set.seed(123) 
fit_nnar_temp <- nnetar(ts_load, xreg = daily_data$daily_temp)

# 3. Forecast the next 31 days using the forecasted July temperatures
forecast_nnar_temp <- forecast(fit_nnar_temp, xreg = future_temp, h = 31)

# 4. Extract predictions and save to CSV
submission$load <- as.numeric(forecast_nnar_temp$mean)
write.csv(submission, "model 15 submission.csv", row.names = FALSE)

print("Saved model 15 submission.csv")
```

## Model 16: Prophet + US Holiday + Multiplicative Seasonality

Score on Kaggle: 0.0964

```{r prophet_with_holidays_and_multiplicative_seasonality, message=FALSE, warning=FALSE}
# 1. Prepare data (reusing df_prophet_holidays from Model 11/12)
df_prophet_multi <- data.frame(
  ds = daily_data$date,
  y = daily_data$daily_avg
)

# 2. Initialize Prophet with multiplicative seasonality and same hyperparameters from Model 12
m_multi <- prophet(
  yearly.seasonality = 15,
  weekly.seasonality = TRUE,
  daily.seasonality = FALSE,
  seasonality.mode = 'multiplicative',
  changepoint.prior.scale = 0.1
)

# Add US holidays
m_multi <- add_country_holidays(m_multi, country_name = 'US')

# 3. Fit and Predict
m_multi <- fit.prophet(m_multi, df_prophet_multi)
future_multi <- make_future_dataframe(m_multi, periods = 31)
forecast_multi <- predict(m_multi, future_multi)

# 4. Extract and Save
prophet_multi_preds <- tail(forecast_multi$yhat, 31)

submission$load <- prophet_multi_preds
write.csv(submission, "model 16 submission.csv", row.names = FALSE)

print("Saved model 16 submission.csv")
```

## Model 17: Prophet + US Holiday + Multiplicative Seasonality + Conditional Seasonality

Score on Kaggle: 0.0963

```{r prophet_with_holidays_and_multiplicative_and_conditional_seasonality, message=FALSE, warning=FALSE}
# 1. Prepare data with summer/other condition flags
df_prophet_multi_cond <- data.frame(
  ds = daily_data$date,
  y = daily_data$daily_avg
)
df_prophet_multi_cond$is_summer <- month(df_prophet_multi_cond$ds) %in% c(6, 7, 8, 9)
df_prophet_multi_cond$is_other <- !df_prophet_multi_cond$is_summer

# 2. Initialize Prophet with multiplicative seasonality + tuned params + no default weekly
m_multi_cond <- prophet(
  yearly.seasonality = 15,
  weekly.seasonality = FALSE,
  daily.seasonality = FALSE,
  seasonality.mode = 'multiplicative',
  changepoint.prior.scale = 0.1
)

# Add US holidays
m_multi_cond <- add_country_holidays(m_multi_cond, country_name = 'US')

# 3. Add conditional weekly seasonalities (summer vs other)
m_multi_cond <- add_seasonality(m_multi_cond, name = 'summer_weekly', period = 7,
                                fourier.order = 3, condition.name = 'is_summer')
m_multi_cond <- add_seasonality(m_multi_cond, name = 'other_weekly',  period = 7,
                                fourier.order = 3, condition.name = 'is_other')

# 4. Fit and Predict
m_multi_cond <- fit.prophet(m_multi_cond, df_prophet_multi_cond)

future_multi_cond <- make_future_dataframe(m_multi_cond, periods = 31)
future_multi_cond$is_summer <- month(future_multi_cond$ds) %in% c(6, 7, 8, 9)
future_multi_cond$is_other <- !future_multi_cond$is_summer

forecast_multi_cond <- predict(m_multi_cond, future_multi_cond)

# 5. Extract and Save
prophet_multi_cond_preds <- tail(forecast_multi_cond$yhat, 31)

submission$load <- prophet_multi_cond_preds
write.csv(submission, "model 17 submission.csv", row.names = FALSE)

print("Saved model 17 submission.csv")
```

## Model 18: Prophet + US Holiday + Multiplicative Seasonality + Temp + Humidity Regressors

Score on Kaggle: 0.1057

```{r prophet_with_holidays_and_multiplicative_seasonality_and_regressors, message=FALSE, warning=FALSE}
# 1. Forecast humidity
humid_ts <- msts(daily_data$daily_humid, seasonal.periods = 365, start = c(2005, 1))
fit_humid <- auto.arima(humid_ts)
forecast_humid <- forecast(fit_humid, h = 31)
future_humid <- as.numeric(forecast_humid$mean)

# 2. Prepare data with both temperature and humidity
df_prophet_weather <- data.frame(
  ds = daily_data$date,
  y = daily_data$daily_avg,
  temp = daily_data$daily_temp,
  humid = daily_data$daily_humid
)

# 3. Initialize Prophet with multiplicative seasonality and same hyperparameters from Model 12
m_weather <- prophet(
  yearly.seasonality = 15,
  weekly.seasonality = TRUE,
  daily.seasonality = FALSE,
  seasonality.mode = 'multiplicative',
  changepoint.prior.scale = 0.1
)

# Add US holidays
m_weather <- add_country_holidays(m_weather, country_name = 'US')

# Add temperature and humidity as extra regressors
m_weather <- add_regressor(m_weather, 'temp',  standardize = TRUE)
m_weather <- add_regressor(m_weather, 'humid', standardize = TRUE)

# 4. Fit and Predict
m_weather <- fit.prophet(m_weather, df_prophet_weather)

future_weather <- make_future_dataframe(m_weather, periods = 31)
future_weather$temp <- c(daily_data$daily_temp,  future_temp) # future_temp from Model 7
future_weather$humid <- c(daily_data$daily_humid, future_humid)

forecast_weather <- predict(m_weather, future_weather)

# 5. Extract and Save
prophet_weather_preds <- tail(forecast_weather$yhat, 31)

submission$load <- prophet_weather_preds
write.csv(submission, "model 18 submission.csv", row.names = FALSE)

print("Saved model 18 submission.csv")
```

## Model 19: Prophet + US Holiday + Multiplicative Seasonality + Conditional Seasonality + Temp + Humidity Regressors

Score on Kaggle: 0.1063

```{r prophet_with_holidays_and_multiplicative_and_conditional_seasonality_and_regressors, message=FALSE, warning=FALSE}
# 1. Prepare data with seasonal condition flags
df_prophet_weather_cond <- data.frame(
  ds = daily_data$date,
  y = daily_data$daily_avg,
  temp  = daily_data$daily_temp,
  humid = daily_data$daily_humid
)
df_prophet_weather_cond$is_summer <- month(df_prophet_weather_cond$ds) %in% c(6, 7, 8, 9)
df_prophet_weather_cond$is_other <- !df_prophet_weather_cond$is_summer

# 2. Initialize Prophet with multiplicative seasonality + tuned params + no default weekly
m_weather_cond <- prophet(
  yearly.seasonality = 15,
  weekly.seasonality = FALSE,
  daily.seasonality = FALSE,
  seasonality.mode = 'multiplicative',
  changepoint.prior.scale = 0.1
)

# Add US holidays
m_weather_cond <- add_country_holidays(m_weather_cond, country_name = 'US')

# Add conditional weekly seasonality
m_weather_cond <- add_seasonality(m_weather_cond, name = 'summer_weekly', period = 7,
                                  fourier.order = 3, condition.name = 'is_summer')
m_weather_cond <- add_seasonality(m_weather_cond, name = 'other_weekly',  period = 7,
                                  fourier.order = 3, condition.name = 'is_other')

# Add weather regressors
m_weather_cond <- add_regressor(m_weather_cond, 'temp',  standardize = TRUE)
m_weather_cond <- add_regressor(m_weather_cond, 'humid', standardize = TRUE)

# 3. Fit and Predict
m_weather_cond <- fit.prophet(m_weather_cond, df_prophet_weather_cond)

future_weather_cond <- make_future_dataframe(m_weather_cond, periods = 31)
future_weather_cond$is_summer <- month(future_weather_cond$ds) %in% c(6, 7, 8, 9)
future_weather_cond$is_other  <- !future_weather_cond$is_summer
future_weather_cond$temp <- c(daily_data$daily_temp,  future_temp)
future_weather_cond$humid <- c(daily_data$daily_humid, future_humid)

forecast_weather_cond <- predict(m_weather_cond, future_weather_cond)

# 4. Extract and Save
prophet_weather_cond_preds <- tail(forecast_weather_cond$yhat, 31)

submission$load <- prophet_weather_cond_preds
write.csv(submission, "model 19 submission.csv", row.names = FALSE)

print("Saved model 19 submission.csv")
```
