---
title: "Forecasting Competition"
author: "Saemin Kim and Matthew Fischer"
output: pdf_document
always_allow_html: true
editor_options: 
  chunk_output_type: console
---

# 1. Data Wrangling and Vanila Model

## Setting R code chunk options

First R code chunk is used for setting the options for all R code chunks. The choice echo=TRUE means both code and output will appear on report, include = FALSE neither code nor output is printed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

## Loading packages and initializing

Second R code chunk is for loading packages. By setting message = FALSE, the code will appear but not the output.

```{r package, message=FALSE}
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)

library(kableExtra)
library(readxl)
library(openxlsx)
```

## Importing data

```{r}
data <- read_excel("load.xlsx")

daily_data <- data %>%
  mutate(
    date = as.Date(date),
    Year = year(date),
    Month = month(date),
    Day = day(date)
  ) %>%
  
  rowwise() %>%
  mutate(
    daily_avg = mean(c_across(h1:h24), na.rm = TRUE)
  ) %>%
  ungroup() %>%
  
  select(meter_id, date, Year, Month, Day, daily_avg)

daily_data <- daily_data %>%
  filter(!is.na(daily_avg), !is.nan(daily_avg), is.finite(daily_avg))
```

```{r}
ggplot(daily_data, aes(x = date, y = daily_avg)) +
  geom_line() +
  labs(title = "Daily Average Load", y = "Daily Average", x = "Date")
```

```{r}
data_ts <- msts(daily_data$daily_avg, 
                            seasonal.periods =c(7, 365),
                            start=c(2005,01))

SNAIVE_data <- snaive(data_ts, h=31)
plot(SNAIVE_data)

forecast_values <- as.numeric(SNAIVE_data$mean)
```

```{r}
submission <- read.csv("submission_template.csv")

submission$load <- forecast_values
write.csv(submission, "seasonal_naive_model.csv", row.names = FALSE)
```

# 2. External Variables and Train/Test Split

## Temperature, Humidity Data Wrangling

```{r external_data, message=FALSE, warning=FALSE}

temp <- read_excel("temperature.xlsx")
humid <- read_excel("relative_humidity.xlsx")

# daliy avg of temp
temp_daily <- temp %>%
  mutate(date = as.Date(date)) %>%
  rowwise() %>%
  # hourly avg of all t_ws1 ~ t_ws28
  mutate(hourly_temp_avg = mean(c_across(starts_with("t_ws")), na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(date) %>%
  # daily avg
  summarise(daily_temp = mean(hourly_temp_avg, na.rm = TRUE))

# daliy avg of humid 
humid_daily <- humid %>%
  mutate(date = as.Date(date)) %>%
  rowwise() %>%
  # hourly avg of all rh_ws1 ~ rh_ws28
  mutate(hourly_humid_avg = mean(c_across(starts_with("rh_ws")), na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(date) %>%
  # daily avg
  summarise(daily_humid = mean(hourly_humid_avg, na.rm = TRUE))

# Join the data
daily_data <- daily_data %>%
  left_join(temp_daily, by = "date") %>%
  left_join(humid_daily, by = "date")

head(daily_data)
```

## Train and Validation Split

```{r train_validation_split, message=FALSE, warning=FALSE}
# Training = 2005-01-01 to 2010-06-30 / Validation = 2010-07-01 to 2010-07-31
train_data <- daily_data %>% filter(date >= as.Date("2005-01-01") & date <= as.Date("2010-06-30"))
valid_data <- daily_data %>% filter(date >= as.Date("2010-07-01") & date <= as.Date("2010-07-31"))

# Create msts (Multi-Seasonal Time Series) objects for training and validation
# Apply weekly (7) and yearly (365) seasonal periods
train_ts <- msts(train_data$daily_avg, seasonal.periods = c(7, 365), start = c(2005, 1))
valid_ts <- msts(valid_data$daily_avg, seasonal.periods = c(7, 365), start = c(2010, 182)) 
# 182 represents the approximate day of the year for July 1st
```

# 3. Top 5 Models

# APPENDIX. Forecast Models and Scores

## Score Summary

| Model Number | Model Description | Score on Kaggle | Rank |
|:-----------------|:-----------------|:-----------------|:-----------------|
| **Model 1** | TBATS | 0.1021 |  **TOP 4**  |
| **Model 2** | STLF-ARIMA | 0.1315 |  |
| **Model 3** | STLF + ETS | 0.1330 |  |
| **Model 4-1** | ARIMA + Fourier terms (K = c(3,10)) | 0.1350 |  |
| **Model 4-2** | ARIMA + Fourier terms (K = c(3,15)) | 1.4840 |  |
| **Model 4-3** | ARIMA + Fourier terms (K = c(2,5)) | 0.1469 |  |
| **Model 5** | NNAR | 0.1770 |  |
| **Model 6** | Ensemble Model (Averaging best outputs) | 0.1211 |  |
| **Model 7** | ARIMAX with temperature forecast | 0.2172 |  |
| **Model 8** | Prophet | 0.0973 | **TOP 3** |
| **Model 9** | Prophet + Temp data | 0.1052 | **TOP 5** |
| **Model 10** | Prophet + ARIMAX | 0.1488 |  |
| **Model 11** | Prophet + US Holiday | 0.0971 | **TOP 2** |
| **Model 12** | Prophet + US Holiday + Parameter tuning | **0.0966** | **TOP 1** |
| **Model 13** | Weighted Ensemble (Tuned Prophet + Base Prophet) | **0.0966** | **TOP 1** |
| **Model 14** | Prophet + US Holiday + Conditional Seasonality | **0.0966** | **TOP 1** |
| **Model 15** | NNAR + Temperature | 0.1287 |  |

## Model 1 : TBATS

Score on Kaggle: 0.1021

```{r model1, message=FALSE, warning=FALSE}
# Create msts object using the FULL dataset (Jan 2005 - Jun 2011)
full_ts <- msts(daily_data$daily_avg, seasonal.periods = c(7, 365), start = c(2005, 1))

# --- Model 1: TBATS Final Forecast ---
print("Retraining TBATS on full data...")
fit_tbats_full <- tbats(full_ts)
forecast_tbats_full <- forecast(fit_tbats_full, h = 31)

# Save TBATS prediction to CSV
submission$load <- as.numeric(forecast_tbats_full$mean)
write.csv(submission, "model 1 submission.csv", row.names = FALSE)
print("Saved model 1 submission.csv")
```

## Model 2: STLF-ARIMA

Score on Kaggle: 0.1315

```{r model2, message=FALSE, warning=FALSE}
# --- Model 2: STLF (with ARIMA) Final Forecast ---
forecast_arima_full <- stlf(full_ts, method = "arima", h = 31)

# Save STLF-ARIMA prediction to CSV
submission$load <- as.numeric(forecast_arima_full$mean)
write.csv(submission, "model 2 submission.csv", row.names = FALSE)
print("Saved model 2 submission.csv")
```

## Model 3: STLF + ETS

Score on Kaggle: 0.1330

```{r model 3, message=FALSE, warning=FALSE}
# --- Model 3: STLF with ETS (Exponential Smoothing) ---
forecast_ets_full <- stlf(full_ts, method = "ets", h = 31)

submission$load <- as.numeric(forecast_ets_full$mean)
write.csv(submission, "model 3 submission.csv", row.names = FALSE)
print("Saved model 3 submission.csv")
```

## Model 4-1: ARIMA + Fourier terms using K = c(3,10)

Score on Kaggle: 0.1350

```{r}
# --- Model 4: Harmonic Regression (ARIMA with Fourier terms) ---
# Using K = c(3, 10)
fit_fourier <- auto.arima(full_ts, seasonal = FALSE, xreg = fourier(full_ts, K = c(3, 10)))
forecast_fourier <- forecast(fit_fourier, xreg = fourier(full_ts, K = c(3, 10), h = 31))

submission$load <- as.numeric(forecast_fourier$mean)
write.csv(submission, "model 4-1 submission.csv", row.names = FALSE)
print("Saved model 4-1 submission.csv")
```

## Model 4-2: ARIMA + Fourier terms using K = c(3,15)

Score on Kaggle: 1.484

```{r}
# --- Model 4: Harmonic Regression (ARIMA with Fourier terms) ---
# Using K = c(3, 15)
fit_fourier <- auto.arima(full_ts, seasonal = FALSE, xreg = fourier(full_ts, K = c(3, 15)))
forecast_fourier <- forecast(fit_fourier, xreg = fourier(full_ts, K = c(3, 15), h = 31))

submission$load <- as.numeric(forecast_fourier$mean)
write.csv(submission, "model 4-2 submission.csv", row.names = FALSE)
print("Saved model 4-2 submission.csv")
```

## Model 4-3: ARIMA + Fourier terms using K = c(2,5)

Score on Kaggle: 0.1469

```{r}
# --- Model 4: Harmonic Regression (ARIMA with Fourier terms) ---
# Using K = c(2, 5)
fit_fourier <- auto.arima(full_ts, seasonal = FALSE, xreg = fourier(full_ts, K = c(2, 5)))
forecast_fourier <- forecast(fit_fourier, xreg = fourier(full_ts, K = c(2, 5), h = 31))

submission$load <- as.numeric(forecast_fourier$mean)
write.csv(submission, "model 4-3 submission.csv", row.names = FALSE)
print("Saved model 4-3 submission.csv")
```

## Model 5: NNAR

Score on Kaggle: 0.1770

```{r}
# --- Model 5: NNAR (Neural Network Time Series Forecast) ---
fit_nnar <- nnetar(full_ts)
forecast_nnar <- forecast(fit_nnar, h = 31)

submission$load <- as.numeric(forecast_nnar$mean)
write.csv(submission, "model 5 submission.csv", row.names = FALSE)
print("Saved model 5 submission.csv")
```

## Model 6: Ensemble Model (Averaging the best outputs)

Score on Kaggle: 0.1211

```{r}
# Averaging the predictions of TBATS, STLF(ARIMA), and STLF(ETS)
ensemble_mean <- (as.numeric(forecast_tbats_full$mean) +
                  as.numeric(forecast_arima_full$mean) + 
                  as.numeric(forecast_ets_full$mean)) / 3

submission$load <- ensemble_mean
write.csv(submission, "model 6 submission.csv", row.names = FALSE)
print("Saved model 6 submission.csv")
```

## Model 7: ARIMAX with temperature forecast

Score on Kaggle: 0.2172

```{r arimax_with_temp_forecast, message=FALSE, warning=FALSE}

# Create a time series object for the historical daily temperature
# Using 365 as the seasonal period since temperature has a strong yearly cycle
temp_ts <- msts(daily_data$daily_temp, seasonal.periods = 365, start = c(2005, 1))

# Fit an ARIMA model to the temperature data
fit_temp <- auto.arima(temp_ts)
# Forecast the temperature for the next 31 days (July 2011)
forecast_temp <- forecast(fit_temp, h = 31)
# Extract the forecasted temperature values as a numeric vector
future_temp <- as.numeric(forecast_temp$mean)


# Forecast the electricity load using the predicted temperatures (ARIMAX)
# Convert full_ts to a simple weekly time series to prevent the auto.arima multi-seasonality error.
# The 'daily_temp' external regressor will naturally capture the yearly (365-day) seasonality.
full_ts_weekly <- ts(daily_data$daily_avg, frequency = 7)

# Fit an ARIMAX model using the weekly ts and historical temperature as an external regressor (xreg)
fit_arimax_full <- auto.arima(full_ts_weekly, xreg = daily_data$daily_temp)

# Forecast the load for July 2011 using the predicted July temperatures as the future xreg
forecast_arimax_final <- forecast(fit_arimax_full, xreg = future_temp, h = 31)

# Save the final predictions to a CSV file for Kaggle submission
submission <- read.csv("submission_template.csv")
submission$load <- as.numeric(forecast_arimax_final$mean)
write.csv(submission, "model 7 submission.csv", row.names = FALSE)
print("Saved model 7 submission.csv")
```

## Model 8: Prophet

Score on Kaggle: 0.0973

```{r prophet_model, message=FALSE, warning=FALSE}
# Load the prophet library
library(prophet)

# 1. Prepare the dataframe in the specific format required by Prophet
df_prophet <- data.frame(ds = daily_data$date, y = daily_data$daily_avg)

# 2. Initialize and fit the Prophet model
# We enable yearly and weekly seasonality. 
# daily.seasonality is set to FALSE because we are working with daily aggregated data.
m <- prophet(yearly.seasonality = TRUE, 
             weekly.seasonality = TRUE, 
             daily.seasonality = FALSE)

m <- fit.prophet(m, df_prophet)

# 3. Create a dataframe for future dates (Next 31 days for July 2011)
future_dates <- make_future_dataframe(m, periods = 31)

# 4. Generate the forecast
forecast_prophet <- predict(m, future_dates)

# Extract only the forecasted values ('yhat') for the last 31 days (July 2011)
prophet_predictions <- tail(forecast_prophet$yhat, 31)

# 5. Save the final predictions
submission <- read.csv("submission_template.csv")
submission$load <- prophet_predictions
write.csv(submission, "model 8 submission.csv", row.names = FALSE)

print("Saved model 8 submission.csv")
```

```{r prophet_visualization, message=FALSE, warning=FALSE}
library(ggplot2)

# 1. Plot the main forecast with historical data and confidence intervals
# Black dots: Actual historical load
# Blue line: Prophet's predicted load
# Light blue shaded area: Uncertainty intervals (Confidence intervals)
print("Plotting the main Prophet forecast...")
plot(m, forecast_prophet) + 
  ggtitle("Prophet Forecast for Daily Load") +
  xlab("Date") + ylab("Daily Average Load")

# 2. Plot the individual components (Trend, Weekly, and Yearly seasonality)
prophet_plot_components(m, forecast_prophet)
```

## Model 9: Prophet + Temp data

Score on Kaggle: 0.1052

```{r prophet_with_regressor, message=FALSE, warning=FALSE}
library(prophet)

print("Preparing data for Prophet with Temperature...")
# 1. Prepare data with both target (y) and the external regressor (temp)
df_prophet_reg <- data.frame(
  ds = daily_data$date, 
  y = daily_data$daily_avg,
  temp = daily_data$daily_temp # Adding historical temperature
)

# 2. Initialize Prophet and ADD the regressor BEFORE fitting
m_reg <- prophet(yearly.seasonality = TRUE, 
                 weekly.seasonality = TRUE, 
                 daily.seasonality = FALSE)
m_reg <- add_regressor(m_reg, 'temp')

print("Fitting Prophet Model with Temperature Regressor...")
m_reg <- fit.prophet(m_reg, df_prophet_reg)

# 3. Create future dataframe
future_dates_reg <- make_future_dataframe(m_reg, periods = 31)
all_temp <- c(daily_data$daily_temp, future_temp)
future_dates_reg$temp <- all_temp

# 5. Generate the final forecast
forecast_prophet_reg <- predict(m_reg, future_dates_reg)

# 6. Extract predictions and save for Kaggle
prophet_reg_predictions <- tail(forecast_prophet_reg$yhat, 31)

submission$load <- prophet_reg_predictions
write.csv(submission, "model 9 submission.csv", row.names = FALSE)
print("Saved model 9 submission.csv")
```

## Model 10: Prophet + ARIMAX

Score on Kaggle: 0.1488

```{r}
ensemble_ultimate <- (prophet_reg_predictions +       
                      as.numeric(forecast_arimax_final$mean)) / 2

submission$load <- ensemble_ultimate
write.csv(submission, "model 10 submission.csv", row.names = FALSE)
print("Saved model 10 submission.csv")
```

## Model 11: Prophet + US Holiday

Score on Kaggle: 0.0971

```{r prophet_with_holidays_only, message=FALSE, warning=FALSE}

# 1. Prepare data
df_prophet_holidays <- data.frame(
  ds = daily_data$date, 
  y = daily_data$daily_avg
)

# 2. Initialize Prophet
m_holidays <- prophet(yearly.seasonality = TRUE, 
                      weekly.seasonality = TRUE, 
                      daily.seasonality = FALSE)

# Add US country holidays (e.g., Independence Day on July 4th)
m_holidays <- add_country_holidays(m_holidays, country_name = 'US')
m_holidays <- fit.prophet(m_holidays, df_prophet_holidays)


# 3. Create future dataframe for July 2011 (31 days)
future_dates_holidays <- make_future_dataframe(m_holidays, periods = 31)

# 4. Generate the final forecast
print("Forecasting with Holidays...")
forecast_prophet_holidays <- predict(m_holidays, future_dates_holidays)

# 5. Extract predictions
prophet_holidays_predictions <- tail(forecast_prophet_holidays$yhat, 31)

submission$load <- prophet_holidays_predictions
write.csv(submission, "model 11 submission.csv", row.names = FALSE)

print("Saved model 11 submission.csv")
```

## Model 12: Prophet + US Holiday + Parameter tuning

Score on Kaggle: 0.0966

```{r prophet_tuned, message=FALSE, warning=FALSE}
# 1. Initialize Prophet with custom hyperparameters
# Increased yearly seasonality flexibility (15 instead of default 10)
# changepoint.prior.scale = 0.1 allows the trend to adapt faster
m_tuned <- prophet(yearly.seasonality = 15, 
                   weekly.seasonality = TRUE, 
                   daily.seasonality = FALSE,
                   changepoint.prior.scale = 0.1)

# Add US holidays
m_tuned <- add_country_holidays(m_tuned, country_name = 'US')

# 2. Fit and Predict
m_tuned <- fit.prophet(m_tuned, df_prophet_holidays)
future_tuned <- make_future_dataframe(m_tuned, periods = 31)
forecast_tuned <- predict(m_tuned, future_tuned)

# 3. Extract and Save
prophet_tuned_preds <- tail(forecast_tuned$yhat, 31)

submission$load <- prophet_tuned_preds
write.csv(submission, "model 12 submission.csv", row.names = FALSE)

print("Saved model 12 submission.csv")
```

## Model 13: Weighted Ensemble (Tuned Prophet + Base Prophet)

Score on Kaggle: 0.0966

```{r final_weighted_ensemble, message=FALSE, warning=FALSE}
# Model A (Rank 1): Tuned Prophet + Holidays (70% weight)
weight_best <- 0.70

# Model B (Rank 2): Base Prophet + Holidays (30% weight)
weight_second <- 0.30

# Calculate the weighted average of the predictions
weighted_ensemble_preds <- (weight_best * prophet_tuned_preds) + 
                           (weight_second * prophet_holidays_predictions)

# Load the submission template and save the final ensemble forecast
submission$load <- weighted_ensemble_preds
write.csv(submission, "model 13 submission.csv", row.names = FALSE)
print("Saved model 13 submission.csv")
```

## Model 14: Prophet + US Holiday + Conditional Seasonality

Score on Kaggle: 0.0966

```{r prophet_conditional, message=FALSE, warning=FALSE}
library(lubridate)

# 1. Prepare data and add conditional columns for Summer vs. Other seasons
df_prophet_ultimate <- data.frame(
  ds = daily_data$date, 
  y = daily_data$daily_avg
)

# Define 'Summer' as June, July, August, and September
df_prophet_ultimate$is_summer <- month(df_prophet_ultimate$ds) %in% c(6, 7, 8, 9)
df_prophet_ultimate$is_other <- !df_prophet_ultimate$is_summer

# 2. Initialize the Prophet model with Tuned Hyperparameters
# Turn OFF default weekly seasonality because we will define custom conditional ones.
m_ultimate <- prophet(yearly.seasonality = 15,          # Tuned parameter
                      weekly.seasonality = FALSE,       # Disabled for custom setup
                      daily.seasonality = FALSE,
                      changepoint.prior.scale = 0.1)    # Tuned parameter

# Add US country holidays
m_ultimate <- add_country_holidays(m_ultimate, country_name = 'US')

# 3. Add Custom Conditional Seasonalities
# This allows the model to learn a different weekly pattern for summer vs other months.
m_ultimate <- add_seasonality(m_ultimate, name = 'summer_weekly', period = 7, 
                              fourier.order = 3, condition.name = 'is_summer')
m_ultimate <- add_seasonality(m_ultimate, name = 'other_weekly', period = 7, 
                              fourier.order = 3, condition.name = 'is_other')
m_ultimate <- fit.prophet(m_ultimate, df_prophet_ultimate)


# 4. Create future dataframe for July 2011 (31 days) and apply conditions
future_ultimate <- make_future_dataframe(m_ultimate, periods = 31)
future_ultimate$is_summer <- month(future_ultimate$ds) %in% c(6, 7, 8, 9)
future_ultimate$is_other <- !future_ultimate$is_summer

# 5. Generate the final forecast
print("Forecasting with Ultimate Prophet...")
forecast_ultimate <- predict(m_ultimate, future_ultimate)

# 6. Extract predictions and save to CSV
prophet_ultimate_preds <- tail(forecast_ultimate$yhat, 31)
submission$load <- prophet_ultimate_preds
write.csv(submission, "model 14 submission.csv", row.names = FALSE)

print("Saved model 14 submission.csv")
```

## Model 15: NNAR + Temperature

Score on Kaggle: 0.1287

```{r nnar_with_temperature, message=FALSE, warning=FALSE}
# Load necessary library
library(forecast)

# 1. Convert historical load to a simple time series object (Weekly frequency)
ts_load <- ts(daily_data$daily_avg, frequency = 7)

# 2. Fit the Neural Network Auto-Regression (NNAR) model
# We use historical daily temperature as an external regressor (xreg)

# The set.seed ensures the neural network initializes the same way every time
set.seed(123) 
fit_nnar_temp <- nnetar(ts_load, xreg = daily_data$daily_temp)

# 3. Forecast the next 31 days using the forecasted July temperatures
forecast_nnar_temp <- forecast(fit_nnar_temp, xreg = future_temp, h = 31)

# 4. Extract predictions and save to CSV
submission$load <- as.numeric(forecast_nnar_temp$mean)
write.csv(submission, "model 15 submission.csv", row.names = FALSE)

print("Saved model 15 submission.csv")
```
